---
title: "Billboard Hot 100 Import."
output:
  html_document:
    df_print: paged
---

```{r setup}
library(tidyverse)
library(janitor)
library(lubridate)
```


## Goals of this notebook

Steps to prepare our data:

- Download the data
- Import into R
- Clean up data types and columns
- Export for next notebook

## Dowloading data

```{r download}
# hot 100 download
download.file("https://github.com/utdata/rwd-billboard-data/blob/main/data-process/hot-100/hot100-orig.csv?raw=true", "data-raw/hot-stuff.csv")
```

## Import data
Importing data from the Billboard Hot 100 data set

```{r import_data}
hot100 <- read_csv("data-raw/hot-stuff.csv") %>% clean_names()

# peek at the data
hot100 %>% glimpse()
```
## Cleanup columns

Use Janitor to make sure our data frames are nice and tidy
## Fix our dates

```{r date_fix}
hot100_date <- hot100 %>% 
  mutate(
    week_id = mdy(week_id)
  ) %>% 
  
  arrange(week_id, week_position)


# peek at the result
hot100_date %>% glimpse()
```

## print the table to look at data

```{r print}
hot100_date %>% head(10)
```
## Selecting columns
Narrowing down the data set to only the columns needed for our analysis

```{r select}
hot100_tight <- hot100_date %>% 
  select(
    -url, 
    -song_id,
    -instance
  )

hot100_tight %>% glimpse()
```

## Exports
Exporting clean data to .rds format

```{r export}
hot100_tight %>% 
  write_rds("data-processed/01-hot100.rds")
```

