---
title: "Billboard analysis"
output:
  html_document:
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })
---

## Exploring the Billboard Top 100 charts dataset

- Who are the 10 performers with the most appearances on the Hot 100 chart at any position?
- Which performer had the most songs reach No. 1?
- Which performer had the most songs reach No. 1 in the most recent five years?
- Which performer had the most Top 10 hits overall?
- Which performer/song combination has been on the charts the most number of weeks at any position?
- Which performer/song combination was No. 1 for the most number of weeks?

## Setup

```{r setup_environment, setup, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
```

## Import data

```{r import_data}
hot100 <- read_rds("data-processed/01-hot100.rds")

# peek at the data
glimpse(hot100)
```

## Analyze Most Appearances

find artist with most appearances on charts

```{r groupby_summarize_appearances}
hot100 %>% 
  group_by(performer) %>% # group by performer
  summarize(
    appearances = n()
    ) %>% # count number of appearances on charts
  arrange(appearances %>% desc()) %>% # arrange results from most to least appearances
  head(10) # show top 10 results
```

### The easier way...

Use count to find the same results

```{r count_appearances}
hot100 %>% 
  count(performer, name ="appearances", sort = TRUE) %>% # count performers
  head(10) # show top 10 results
```

## Most songs to reach No. 1

use `group_by` and `summarize` to find artists with most songs to hit number 1

```{r filter}
hot100 %>% 
  filter(week_position == 1) %>% # look for what hit number 1
  distinct(song, performer) %>% # get unique entries
  group_by(performer) %>% # group results by performer
  summarise(no1hits = n()) %>% #count number of rows in groups, name new columns
  arrange (no1hits %>% desc()) %>% # sorts from highest to lowest
  filter(no1hits >= 7) #limit results to those with 7 or more hits
```

### Using count

find same results using `count` now

```{r}
hot100 %>% 
  filter(week_position == 1) %>% #filter to number 1 hits
  distinct(song, performer) %>% # get unique entries
  count(performer, name = "no1_hits", sort = TRUE) %>% # count by performer, sort top to bottom
  filter(no1_hits >= 7) # limit to 7 or more
```

## No. 1 hits in last 5 years

Find artists with the most number 1 hits in the last 5 years

```{r}
hot100 %>% 
  filter(
    week_position == 1,
    week_id > "2015-12-31"
  ) %>% #filter by number 1 hits of last 5 years
  distinct(song, performer) %>%  # narrow to distinct entries
  count(performer, name = "no1_hits", sort = TRUE) %>% # count number of entries
  filter(no1_hits > 1) #filter to those that appear more than once
```

### Now use `summarize()` and `group_by()`

```{r groupby_and_summarize} 
hot100 %>% 
  filter(
    week_position == 1,
    week_id > "2015-12-31"
  ) %>% # filter by number 1 hits in last 5 years
  distinct(performer, song) %>% # get unique entries
  group_by(performer) %>% # group results by performer
  summarise(no1_hits = n()) %>% # count each entries
  filter(no1_hits > 1) %>% # filter by those that appear more than once
  arrange(no1_hits %>% desc()) # sort most to least
```

## Top 10

find what performers had the most top 10 hits over all time

```{r}
hot100 %>% 
  filter(
    week_position <= 10
  ) %>% # filter for top 10 positions
  distinct(song, performer) %>% # get unique entries
  group_by(performer) %>% # group results by performer
  summarize(number_top_hits = n()) %>% # summarize the amount of times each artist appeared on top ten
  arrange(number_top_hits %>% desc()) %>% # sort by most to least appearances
  filter(number_top_hits > 19) #limit the results logically
```

## Most appearances

Finding which performer/song combination appeared on the charts the most number of weeks overall

```{r}
hot100 %>% 
  count(song, performer, name = "weeks_on_charts", sort = TRUE) %>% # count times performer/song grouping appears
  filter(weeks_on_charts > 63) # limit results
```

## Most weeks at No. 1

Filter for the songs that spent the most weeks at number 1

```{r}
hot100 %>% 
  filter(
    week_position == 1 # get number 1 hits
  ) %>% 
  count(song, performer, name = "weeks_at_number_one", sort = TRUE) %>% # count times songs appeared at the position
  filter(weeks_at_number_one > 13) # limit results
```

## Dig around a bit more...

```{r}
hot100 %>% 
  filter(
    week_position == 1 # get number 1 hits
  ) %>% 
  count(song, performer, name = "weeks_at_number_one", sort = TRUE) %>% # count times songs appeared at the position
  filter(weeks_at_number_one > 13) # limit results
```
